---
title: "OONI Web Connectivity Test Data Exploration"
output:
  html_document:
    df_print: paged
---

```{r warning=FALSE, include=FALSE}
library(tidyverse)
library(knitr)

library(lubridate)
```

## Data Import
Data for OONI Meeasurements of the Web Connectivity Test in the UK from January 2020

```{r echo=TRUE, warning=FALSE}
df <- read.csv('OONI January 2020.csv', na.strings = c('Na','NA', ''))

# row numbers...
df$X <- NULL

df %>% head()
```


## Data Cleaning

```{r}
# transform booleans to loical
df$anomaly <- as.logical(df$anomaly)
df$confirmed <- as.logical(df$confirmed)
df$failure <- as.logical(df$failure)
df$https <- as.logical(df$https)
df$headers_match <- as.logical(df$headers_match)
df$body_length_match <- as.logical(df$body_length_match)

df$measurement_start_time <- as.Date(df$measurement_start_time)
```


### False Anomaly Cloudflare


'Attention Required | Cloudflare', sounds like a Error page from the Cloudflare DNS...


```{r message=FALSE, warning=FALSE, paged.print=TRUE}
df %>% filter(anomaly,grepl("Cloudflare", page_title))
```

### False Pandora detection

pandora.com displays a page that its service is only available from the US.

```{r message=FALSE, warning=FALSE}
df %>% filter(grepl("pandora", input)) %>% group_by(asn_network, dns_network, blocking) %>% count()

```

```{r}
# Yahoo Mail blocked

df %>% filter(anomaly) %>% group_by(input) %>% count() %>% arrange(-n)

df %>% filter(anomaly, input == 'https://mail.yahoo.com/')
```



```{r}

print(paste("Before Cleaning: ",nrow(df), "rows",sum(df$anomaly), "anomalies", sum(df$blocking != "False" & !is.na(df$blocking)), "blocks"))

# Exclude measurements with control probe errors
df <- df %>% 
  filter(control_status_code == 200 | is.na(control_status_code), 
         is.na(control_dns_failure), 
         is.na(control_http_failure), 
         is.na(control_failure), 
         !grepl("pandora", input),
         !grepl("Attention Required! | Cloudflare", page_title))

# dected block in blocking is not NA and not False
df$blocked <- !is.na(df$blocking) & df$blocking != 'False'


print(paste("After Cleaning: ",nrow(df), "rows",sum(df$anomaly), "anomalies", sum(df$blocking != "False" & !is.na(df$blocking)), "blocks"))

```


## Data Exploration

```{r}
# Display column properties
data.frame(Type = sapply(sapply(df, class), `[[`, 1), Dist.Values = sapply(sapply(df, unique), length), NAs = colSums(is.na(df))) %>% kable()

```



### Categorical Values Distributions

```{r}
# Plot values distributions for categorical variables with between 2 and 30 distinct values
for(x in colnames(
  df[ , sapply(sapply(df, unique), length) > 1 
      & sapply(sapply(df, unique), length) < 30 
      & !sapply(df, is.numeric)])){
  
  
  p <- df %>% ggplot(mapping = aes(x = get(x), fill= get(x))) +
  geom_bar()+
  labs(
    title = paste(x, " - ",class(df[ , x])),
    x = ''
  )  +
  geom_text(
     aes(label=paste0(round(stat(prop),3)*100,'%'), group=1),
     stat='count',
     nudge_y = 10
 ) +
  theme_minimal() +
  theme (
    axis.text.x = element_text(angle = 18, hjust = 1),
    legend.position = 'none'    
         )
  
  
  print(p)
  
}
```


### Tested URLs
```{r message=FALSE, warning=FALSE}
# URLs
length(unique(df$input))
# domains
length(unique(df$domain))

# Are all URLs tested by all probes?
df %>% group_by(asn_network) %>% summarise(total = n(),unique_urls =length(unique(domain)) ) %>% arrange(-total)

# Seems that URLs are not tested equally often by all probes

```

### Which Domains most often Blocked?

```{r}
# How many measurements are blocked ?
nrow(df %>% filter(blocked)) / nrow(df)


# Most tested ISPs
df %>% 
  filter(blocked) %>%
  group_by(domain, page_title) %>% 
  count() %>% 
  arrange(-n)
```


## Most common HTTP response status codes

```{r message=FALSE, warning=FALSE}
df %>% 
  group_by(http_status_code) %>% 
  count() %>% 
  mutate(total_share = (n / nrow(df)) ) %>%
  arrange(-n)
```

https://restfulapi.net/http-status-codes/

200 - ok
301 - moved permamently
302 - found URL redirection
307 - temporarly redirect
403 - Forbidden
404 - Not Found
500 - Server Error


### Any Legal HTTP blocks?
```{r}
df %>% filter(grepl("451", http_status_code))

```

3 cases with a reported legal Block...
Exponential-E Ltd.



### Blocking Mechanisms
```{r}
df %>% 
  filter(blocked) %>%
  group_by(blocking) %>%
  count()
```


### Time Series

```{r}
# get the dat from measurement start
df <- df %>% mutate(dt = date(measurement_start_time))

# How many measurements per day??
df %>% group_by(dt) %>% count() %>% ggplot(aes(x = dt, y = n)) + geom_bar(stat = 'sum') + ggtitle("Measurements per Day") 

# How many distinct domains tested that day ?
df %>% group_by(dt) %>% summarise(domains_tested = length(unique(input))) %>% ggplot(aes(x = dt, y = domains_tested)) + geom_bar(stat = 'sum') + ggtitle("URLs tested per Day")

 
```


## Comparison blocking per ISP

### Non Mobile Networks

```{r warning=FALSE}
df %>% 
  filter(probe_network_type != 'mobile' | is.na(probe_network_type) ) %>%
  group_by(asn_network) %>% 
  summarise(Measurements= n(),  URLs = length(unique(input)), blocked_urls= length(unique(input[blocked]))) %>% 
  arrange(-Measurements) #%>%
  #write.csv('non Mobile ISPs.csv', row.names = FALSE)
```

### Mobile Networks

```{r warning=FALSE}
df %>% 
  filter(probe_network_type == 'mobile') %>%
  group_by(asn_network) %>% 
  summarise(Measurements= n(),  URLs = length(unique(input)), blocked_urls= length(unique(input[blocked]))) %>% 
  arrange(-Measurements) #%>%
  #write.csv('Mobile ISPs.csv', row.names = FALSE)

```



## Blocking Techniques per ISP

### All Networks 

```{r warning=FALSE}
df %>% 
  filter(blocked) %>% 
  count(asn_network, blocking) %>% 
  spread(blocking,n) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(Total = rowSums(.[2:5])) %>%
  arrange(-Total) %>%
  select(asn_network, Total, dns,'http-diff', 'http-failure', tcp_ip)

```


### HTTP vs HTTPS blockings

```{r warning=FALSE}
# How many distinct URLs blocked
df %>% 
  group_by(https) %>% 
  summarise(Measurements= n(),  Blocked = sum(blocked), URLs = length(unique(input)), blocked_urls= length(unique(input[blocked]))) %>% 
  mutate(Share_blocked = Blocked / Measurements, Share_blocked_URLs = blocked_urls / URLs) %>%
  arrange(-Measurements) 

```



```{r warning=FALSE}
# HTTP 
df %>% 
  filter(!https) %>% 
  count(asn_network, blocking) %>% 
  spread(blocking,n) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(Total = rowSums(.[2:7]) , Blocked = rowSums(.[ c('dns','http-diff', 'http-failure', 'tcp_ip') ]), blocked_share = Blocked / Total ) %>%
  select(asn_network, Total, Blocked, blocked_share, dns, 'http-failure', 'http-diff', tcp_ip) %>%
  arrange(-Total) #%>%
  #write.csv('blocking_technique HTTP.csv', row.names = FALSE)

# HTTPS
df %>% 
  filter(!https) %>% 
  count(asn_network, blocking) %>% 
  spread(blocking,n) %>% 
  mutate_all(~replace(., is.na(.), 0)) %>% 
  mutate(Total = rowSums(.[2:7]) , Blocked = rowSums(.[ c('dns','http-diff', 'http-failure', 'tcp_ip') ]), blocked_share = Blocked / Total ) %>%
  select(asn_network, Total, Blocked, blocked_share, dns, 'http-failure', 'http-diff', tcp_ip) %>%
  arrange(-Total) #%>%
  #write.csv('blocking_technique HTTP.csv', row.names = FALSE)

```


### DNS blocking 
```{r warning=FALSE}
df %>%
  group_by(asn_network,dns_network) %>%
  summarise(Measurements = n(), Blocks = sum(blocked), DNS_Blocks = sum(blocked[blocking == 'dns'])) %>%
  arrange(-Measurements) 

```


## Age Verification Pages 

```{r}
df %>% 
  filter(blocked, probe_network_type == 'mobile', blocking =='http-diff', asn_network =='Telefonica UK Limited') 
  
```

https://api.ooni.io/api/v1/measurement/temp-id-377435600 depicts a age verification page from O2


